Symbolic Abstract Data Type Inference

### 2015-07-09 overview.tex
Proper algorithm rendering.

Our basic approach to inferring ADT specifications, as outlined by the abstract
algorithm in Figure~\ref{fig:abstract}, is to identify a finite set of
sequential execution histories which capture all of the reasons for which a
sequence could be considered invalid, according to the ADT of a given reference
implementation. These sequences thus serve as patterns indicating violations in
the sequences which contain them. Thus the linearizations which exclude all
violation patterns are considered valid.

\begin{figure}[t]
  \begin{verbatim}
procedure INFER_ADT_FROM_REFERENCE_IMPL(I)
  PATTERNS := {}
  for each sequential history H
    if * then
      break
    else if H is EXECUTABLE with I then
      continue
    else if H is REDUNDANT with PATTERNS then
      continue
    else
      add H to PATTERNS
    end
  end
  return EXCLUSION of PATTERNS
end
  \end{verbatim}
  \caption{An abstract algorithm for symbolic ADT inference.}
  \label{fig:abstract}
\end{figure}

@@

The key technical obstacle in realizing the abstract algorithm of
Figure~\ref{fig:abstract}, overcome in Section~\ref{sec:patterns}, is the
classification of violations into a finite set of patterns, rendering the
remaining invalid sequences redundant. The symbolic representation which
excludes a given set of patterns is relatively straightforward to construct, and
is given in Section~\ref{sec:formula}. Note however that the termination of this
abstract algorithm is nondeterministic, and thus the inferred ADT specification
is generally not the strongest possible in the sense that it may not exclude
certain invalid linearizations. Nevertheless, specifications inferred by this
algorithm are sound, in the sense that when modular program reasoning with
inferred specifications succeeds, correctness follows. In any case, the
specifications inferred by our refinement of this abstract algorithm in
Section~\ref{sec:algorithm} are both sound and complete for the
naturally-occurring ADTs of interest. The remainder of this article develops
the technical machinery required to realize this abstract inference algorithm.

>>>>

Our basic approach to inferring ADT specifications, as outlined by the abstract
algorithm in Algorithm~\ref{alg:abstract}, is to identify a finite set of
sequential execution histories which capture all of the reasons for which a
sequence could be considered invalid, according to the ADT of a given reference
implementation. These sequences thus serve as patterns indicating violations in
the sequences which contain them. Thus the linearizations which exclude all
violation patterns are considered valid.

\begin{algorithm}[t]
  \SetAlgoLined
  \SetKwInOut{Input}{input}
  \Input{A reference implementation $\mathcal{I}$}
  \KwResult{A formula representing the ADT of $\mathcal{I}$}
  patterns $\gets$ $\emptyset$ \;
  \For{each sequential history $h$}{
    \uIf{$*$}{
      break
    }
    \uElseIf{$h$ is \emph{executable} with $\mathcal{I}$}{
      continue
    }
    \uElseIf{$h$ is \emph{redundant} with patterns}{
      continue
    }
    \Else{
      add $h$ to patterns
    }
  }
  \Return \emph{exclusion} of patterns
  \caption{Abstract algorithm for symbolic ADT inference.}
  \label{alg:abstract}
\end{algorithm}

@@

The key technical obstacle in realizing the abstract algorithm of
Algorithm~\ref{alg:abstract}, overcome in Section~\ref{sec:patterns}, is the
classification of violations into a finite set of patterns, rendering the
remaining invalid sequences redundant. The symbolic representation which
excludes a given set of patterns is relatively straightforward to construct, and
is given in Section~\ref{sec:formula}. Note however that the termination of this
abstract algorithm is nondeterministic, and thus the inferred ADT specification
is generally not the strongest possible in the sense that it may not exclude
certain invalid linearizations. Nevertheless, specifications inferred by this
algorithm are sound, in the sense that when modular program reasoning with
inferred specifications succeeds, correctness follows. In any case, the
specifications inferred by our refinement of this abstract algorithm in
Section~\ref{sec:algorithm} are both sound and complete for the
naturally-occurring ADTs of interest. The remainder of this article develops
the technical machinery required to realize this abstract inference algorithm.
